---
title: "Gender and social media use"
author: "Dipankar Lahiri"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
library (dplyr)
library (readr)
library (ggplot2)
library (rcompanion)

my_df <- read_csv ('BES-2017-f2f-gender-fb-tw.csv')

```


# This is an analysis of a random sample dataframe of 2189 individuals classified on Gender into Male and Female. Their responses on use of Facebook and Twitter respectively were entered as binary nominal variables. 

# Main findings
# 1. Men are 31% less likely than women to use Facebook. 
# 2. Women are 23.5% less likely than men to use Twitter.
# 3. 14.8% (95% CI: 13.3% - 16.3%) people use both Facebook and Twitter. 40.9% (95% CI: 38.9% - 42.9%) people use neither.
# 4. Those who use Twitter are more than 7 times more likely to be Facebook users as well than those who do not use Twitter. 
# 5. The likeliest person in our sample to be a Facebook user is a Twitter-using Female (90.6% predicted probability) and the least likely person to be a Facebook user is a non-Twitter-using Male (45% predicted probability).


```{r summarising variables}

my_df %>% count(Gender) %>% mutate(percentage = n / sum(n) * 100)
my_df %>% count(Use_Facebook) %>% mutate(percentage = n / sum(n) * 100)
my_df %>% count(Use_Twitter) %>% mutate(percentage = n / sum(n) * 100)

```

# Summary of variable counts: The sample had 1000 males (46%) and 1189 females (54%). The sample had 1251 Facebook users and 938 non-users, or 57% of the sample were Facebook users. The sample had 366 Twitter users and 1823 non-users, or 17% of the sample were Twitter users.


# Hypotheses: There is an association between gender and social media use.


# GENDER AND FACEBOOK

```{r hypothesis_test_FB}

chisq.test(
  x = my_df$Gender,
  y = my_df$Use_Facebook
)

```


```{r visualisation}

my_df %>%
  ggplot (aes(x=Gender, fill = Use_Facebook)) + geom_bar (position = 'dodge')


```

# GENDER AND TWITTER

```{r hypothesis_test_Twitter}
chisq.test(
  x = my_df$Gender,
  y = my_df$Use_Twitter
)
```


```{r}
my_df %>%
  ggplot (aes(x=Gender, fill = Use_Twitter)) + geom_bar (position = 'dodge')

```

# Findings - A Pearson’s Chi-Square test of independence shows a significant positive relationship between Gender and Facebook use, χ²(1, N = 2189) = 18.05, p < 0.001. 

# On the other hand, a Pearson’s Chi-Square test shows no significant relationship between Gender and Twitter use, χ²(1, N = 2189) = 0.705, p = 0.401 (p>0.05). This could be because of the low proportion of Twitter users in the sample (17%).


# TWITTER USE AND FACEBOOK USE

```{r}
chisq.test(
  x= my_df$Use_Facebook,
  y= my_df$Use_Twitter
)

my_df %>% ggplot(aes(x = Use_Facebook, fill = Use_Twitter)) + geom_bar (position = 'dodge')

```


# A Pearson’s Chi-Square test of independence shows a significant positive relationship between Facebook use and Twitter use, χ²(1, N = 2189) = 175.13, p < 0.001. 

# We see that Gender and Twitter use both have significant relationships with Facebook use. The Twitter-Facebook relationship seems to be stronger than the Gender-Facebook relationship based on a higher X² value returned in the Chi-Square test. Let's explore these two relationships further with Fisher's exact test for count data. 


```{r}
fisher.test(
  x = my_df$Gender,
  y = my_df$Use_Facebook
)

fisher.test(
  x = my_df$Use_Twitter,
  y = my_df$Use_Facebook
)
```

# Findings: A Fisher's exact test shows a significant negative relationship between Gender and Facebook use, OR = 0.69, 95% CI [0.58, 0.82], p < 0.001, suggesting that men are 31% less likely than women to use Facebook. The 95% confidence interval indicates that the true population odds ratio is likely between 0.58 and 0.82, meaning men’s Facebook use is at least 18% lower and at most 42% lower than that of women.

# A Fisher's exact test shows a significant positive relationship between Twitter use and Facebook use, OR = 7.45, 95% CI [5.31, 10.68], p < 0.001, suggesting that those who use Twitter are more than 7 times more likely to use Facebook compared to those who who do not use Twitter. The 95% confidence interval indicates that the true population odds ratio is likely between 5.31 and 10.68, meaning that Twitter users have at least 5.3 times higher odds and at most 10.7 times higher odds of using Facebook than non-Twitter users.

# LOGISTIC REGRESSION

```{r}
my_df <- my_df %>%
  mutate(
    Gender = factor(Gender),
    Use_Facebook = factor(Use_Facebook, levels = c("No", "Yes")),
    Use_Twitter = factor(Use_Twitter, levels = c("No", "Yes")) 
  )

model <- glm(Use_Facebook ~ Gender + Use_Twitter, data = my_df, family = binomial)

summary(model)

my_df$Predicted_Prob <- predict(model, type = "response")
head(my_df[, c("Gender", "Use_Twitter", "Predicted_Prob")])

new_data <- data.frame(
  Gender = factor(c("Male", "Male", "Female", "Female"), levels = c("Male", "Female")),
  Use_Twitter = factor(c("No", "Yes", "No", "Yes"), levels = c("No", "Yes"))
)

new_data$Predicted_Prob <- predict(model, newdata = new_data, type = "response")

new_data

ggplot(new_data, aes(x = interaction(Gender, Use_Twitter), y = Predicted_Prob, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Gender & Twitter Use", y = "Predicted Probability of Facebook Use", title = "Predicted Probabilities of Facebook Use") +
  theme_minimal()

ggplot(new_data, aes(x = interaction(Gender, Use_Twitter), y = Predicted_Prob, fill = Gender)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = round(Predicted_Prob, 2)), vjust = -0.5) +
  labs(x = "Gender & Twitter Use", y = "Predicted Probability of Facebook Use", title = "Predicted Probabilities of Facebook Use") +
  theme_minimal()

ggplot(new_data, aes(x = Gender, y = Use_Twitter, fill = Predicted_Prob)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "Gender", y = "Twitter Use", fill = "Probability", title = "Predicted Probabilities of Facebook Use") +
  theme_minimal()

knitr::kable(exp(cbind(OR = coef(model), confint.default(model))))


```

## Process: A binary logistic regression model was conducted to predict Facebook usage based on Gender and Twitter usage. The model was statistically significant, χ²(2, N = 2189) = 224.7, p < 0.001, indicating that the predictors explained a significant portion of variance in Facebook usage. The Akaike Information Criterion (AIC) was 2771, suggesting an improved model fit compared to the null model (AIC = 2989.7).

## Findings
# The likeliest person in our sample to be a Facebook user is a Twitter-using Female (90.6% predicted probability) and the least likely person to be a Facebook user is a non-Twitter-using Male (45% predicted probability).
# Gender: Men were significantly less likely to use Facebook than women (β = -0.431, p < 0.001). The odds of Facebook use for men were 35% lower compared to women (OR = 0.65, 95% CI: 0.54 - 0.78). The predicted probability of Facebook use was ~55.7% for women and ~44.9% for men, holding Twitter usage constant. 
# Twitter Usage: Those who used Twitter were significantly more likely to use Facebook (β = 2.038, p < 0.001). Twitter users had 7.67 times higher odds of using Facebook compared to non-Twitter users (OR = 7.67, 95% CI: 5.48 - 10.73). 
# Twitter users are more likely to use Facebook even after controlling for Gender. Gender also has an independent effect on Facebook use.

# PREDICTING GENDER FROM SOCIAL MEDIA USE

```{r}
N <- 2189
z <- 1.96
both_count <- my_df %>% filter(Use_Twitter == "Yes" & Use_Facebook == "Yes") %>% nrow()
neither_count <- my_df %>% filter(Use_Twitter == "No" & Use_Facebook == "No") %>% nrow()

p_both <- both_count / N
p_neither <- neither_count / N

se <- function(p, N) sqrt((p * (1 - p)) / N)

se_both <- se(p_both, N)
se_neither <- se(p_neither, N)

ci_both <- c(p_both - z * se_both, p_both + z * se_both)
ci_neither <- c(p_neither - z * se_neither, p_neither + z * se_neither)

my_df$Gender_binary <- ifelse(my_df$Gender == "Female", 1, 0)

logit_model <- glm(Gender_binary ~ Use_Twitter + Use_Facebook, data = my_df, family = binomial)

summary(logit_model)
exp(coef(logit_model))
exp(confint(logit_model))

new_data <- data.frame(Use_Twitter = c("Yes", "Yes", "No", "No"), Use_Facebook = c("Yes", "No", "Yes", "No"))

new_data$predicted_female <- predict(logit_model, newdata = new_data, type = "response")

new_data

```

# Proportion using both Twitter and Facebook: 14.8% (95% CI: 13.3% - 16.3%). Proportion using neither platform: 40.9% (95% CI: 38.9% - 42.9%). 

# A binary logistic regression was conducted to predict Gender based on Twitter and Facebook usage. Women were 23.5% less likely than men to use Twitter (OR = 0.77, 95% CI: 0.60 - 0.97, p < 0.001). Women were 53.8% more likely than men to use Facebook (OR = 1.54, 95% CI: 1.29 - 1.84, p < 0.001). This suggests Facebook-only users have the highest probability of being Female, while Twitter-only users have the highest probability of being Male. Those who use neither platform are closer to an even gender split.