---
title: "Markdown"
author: "Dipankar Lahiri"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
install.packages('rcompanion')
library (dplyr)
library (readr)
library (ggplot2)
library (rcompanion)

my_df <- read_csv ('BES-2017-f2f-gender-fb-tw.csv')

```

#H1: There is an association between gender and social media use.

``` {r hypothesis test - FB}

chisq.test(
  x = my_df$Gender,
  y = my_df$Use_Facebook
)

```


``` {r visualisation}

my_df %>%
  ggplot (aes(x=Gender, fill = Use_Facebook)) + geom_bar (position = 'dodge')


```

# Homework: Gender and Twitter

``` {r hypothesis test - Twitter}
chisq.test(
  x = my_df$Gender,
  y = my_df$Use_Twitter
)
```


``` {r}
my_df %>%
  ggplot (aes(x=Gender, fill = Use_Twitter)) + geom_bar (position = 'dodge')

```

## Findings - A Pearson’s Chi-Square test of independence shows a significant positive relationship between Gender and Facebook usage, χ²(1, N = 2189) = 18.05, p < 0.001. On the other hand, a Pearson’s Chi-Square test shows no significant relationship between Gender and Twitter usage, χ²(1, N = 2189) = 0.705, p = 0.401 (p>0.05).

# Twitter and Facebook usage

```{r}
chisq.test(
  x= my_df$Use_Facebook,
  y= my_df$Use_Twitter
)

my_df %>% ggplot(aes(x = Use_Facebook, fill = Use_Twitter)) + geom_bar (position = 'dodge')

```
# We know now that Gender and Twitter use both have significant relationships with Facebook use. Which has a greater association? 

```{r}
fisher.test(
  x = my_df$Gender,
  y = my_df$Use_Facebook
)

fisher.test(
  x = my_df$Use_Twitter,
  y = my_df$Use_Facebook
)
```

# Findings: A Fisher's exact test shows a significant negative relationship between Gender and Facebook use, OR = 0.69, 95% CI [0.58, 0.82], p < 0.001, suggesting that men are 31% less likely than women to use Facebook. A Fisher's exact test shows a significant positive relationship between Twitter use and Facebook use, OR = 7.45, 95% CI [5.31, 10.68], p < 0.001, suggesting that those who use Twitter are more than 7 times more likely to use Facebook as well.

```{r}



